---
title: "Programming Languages für Data Science"
subtitle: "Major Assignment: Flight Analysis - Data Preparation and first Insights"
author: "Christian Roth & Andreas Waßmus"
date: "22.10.2019"
output:
    html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 Data Preparation


## 1.1 Libraries and Data

```{r}
library(tidyverse)
library(readr)
library(DataExplorer)
```



Loading the data:

```{r}
usflights <- read_csv("./data/flight-delays/usflights15.csv")
```


```{r}
nrow(usflights)
```

The data contains 5.819.079 datasets (=count of flights in the US in 2015)


```{r}
head(usflights)
```

Important features for analysis:
Date
Airline
Flight_Number
Origin_Airport
Destination_Arport
Scheduled_Departure
Departure Time
Departure Delay


## 1.2 Missing values in the flight data

```{r}
plot_missing(usflights)
```

You can also count the missing value per column with this function:

```{r}
#is.na(usflights)
colSums(is.na(usflights))
```


In the dataset there is a high amount of missing values in the features: CANCELLATION_REASON, AIR_SYSTEM_DELAY, SECURITY_DELAY, AIRLINE_DELAY, LATE_AIRCRAFT_DELAY, WEATHER_DELAY. Since these colums are only explanations of the features cancellation and delay we can remove them from the dataset.


```{r}
usflights$CANCELLATION_REASON <- NULL
usflights$AIR_SYSTEM_DELAY <- NULL
usflights$SECURITY_DELAY <- NULL
usflights$AIRLINE_DELAY <- NULL
usflights$LATE_AIRCRAFT_DELAY <- NULL
usflights$WEATHER_DELAY <- NULL
head(usflights, 3)
```


```{r}
colSums(is.na(usflights))
```

For the analysis it is for us important to have a complete datasets. Therefore, rows with missing values will be deleted to have only complete observations. (ARRIVAL_DELAY could be also calculated by "mutate". For simplification it is waived due to the large number of records.)

```{r}
usflights_complete <- usflights[!is.na(usflights$DEPARTURE_TIME),]
usflights_complete <- usflights[!is.na(usflights$ARRIVAL_TIME),]
usflights_complete <- usflights[!is.na(usflights$ARRIVAL_DELAY),]
```

Check if data is complete and without missing values:

```{r}
colSums(is.na(usflights_complete))
```

```{r}
nrow(usflights_complete)
```

105.071 data sets were reomved from the analyse. The data has now 5.714.008 oberservations.


## 1.3 Merge flight data with arilines und airports

Read csv and save as data frame:

```{r}
airlines2015 <- read_csv("./data/flight-delays/airlines15.csv")
airports2015 <- read_csv("./data/flight-delays/airports15.csv")
```

Overview of the data:

```{r}
data_list_2015 <- list(airlines2015, airports2015, usflights)
plot_str(data_list_2015)
```



Merge df airlines2015 to flight dataset:

```{r}
flights2015 <- merge(usflights_complete, airlines2015, by.x= "AIRLINE", by.y="IATA_CODE", all.x = TRUE)

```

Merge df airports2015 to flight dataset:

```{r}
flights2015_merged <- merge(flights2015, airports2015, by.x = "ORIGIN_AIRPORT", by.y = "IATA_CODE", all.x = TRUE)
```

Check if merge has worked successfully:

```{r}
head(flights2015_merged)
```


## 1.4 Missing values in the merged data

Check missing values per column:

```{r}
colSums(is.na(flights2015_merged))
```

```{r}
plot_missing(flights2015_merged)
```

As bevor mentioned, it is important for the analysis to have complete datasets. Therefore, the above shown features with missing values will be deleted.

```{r}
flights2015_merged_complete <- flights2015_merged[!is.na(flights2015_merged$COUNTRY),]
flights2015_merged_complete <- flights2015_merged[!is.na(flights2015_merged$STATE),]
flights2015_merged_complete <- flights2015_merged[!is.na(flights2015_merged$CITY),]
flights2015_merged_complete <- flights2015_merged[!is.na(flights2015_merged$AIRPORT),]
flights2015_merged_complete <- flights2015_merged[!is.na(flights2015_merged$LONGITUDE),]
flights2015_merged_complete <- flights2015_merged[!is.na(flights2015_merged$LATITUDE),]
```


Check if all rows with missing value are excluded from the analysis.

```{r}
colSums(is.na(flights2015_merged_complete))
```

```{r}
nrow(flights2015_merged_complete)
```

The data has 5.226.569 complete oberservations. Starting from 5.819.079 rows, this represants a reduction of 592.510 or 10,2%. 

```{r}
nrow(flights2015_merged)
```


## 1.5 Building DEPARTURE_DATE From the columns YEAR, MONTH and DAY.

This step makes our analysis easier and is important for showing flight movements in a map for each day.

```{r}
flights2015_merged_complete$DEPARTURE_DATE <- as.Date(with(flights2015_merged_complete, paste(YEAR, MONTH, DAY, sep = "-")),"%Y-%m-%d")
head(flights2015_merged_complete$DEPARTURE_DATE)
```

## 1.6 Creation of 3 data frames and save as .csv

### 1.6.1 Complete data

For further processing we save a csv from this dataframe:

```{r}
write.csv(flights2015_merged_complete,("./data/cleandata/flights15_complete.csv"))
```


```{r}
flights15_complete_sample <- sample_n(flights2015_merged_complete, 50)
write.csv(flights15_complete_sample,("./data/cleandata/flights15_complete_sample.csv"))

```


### 1.6.2 Christmas data 2015 (23.12.-26.12.)

In fact it is very hard to work with such a big dataframe and every analysis will take a lot of time. Therefore, we decided to conduct a Christmas Flight Study in 2015 and we will work with a smaler data frame:


```{r}
christmas15 <- filter(flights2015_merged_complete, DEPARTURE_DATE >="2015-12-23" & DEPARTURE_DATE <= "2015-12-26")
```

56.246 oberservations were selected for the further analysis:

```{r}
nrow(christmas15)
```

```{r}
#write.csv(christmas15,("./data/cleandata/christmas15.csv"))
```

### 1.6.3 Christmas Day data (25.12.)

Dataset for an animated flight map:

```{r}
christmasDay15 <- filter(christmas15, DEPARTURE_DATE == "2015-12-25" & ORIGIN_AIRPORT == "JFK" | DESTINATION_AIRPORT == "JFK")
head(christmasDay15)
```

```{r}
nrow(christmasDay15)
```

## 1.6.4 Random Sample of flights at Christmas Day 

Random sample fo December, 25th:

```{r}
christmasDay15_sample <- sample_n(christmasDay15, 50)
head(christmasDay15_sample)
```

```{r}
#write.csv(christmasDay15,("./data/cleandata/christmasDay15.csv"))
```

```{r}
#write.csv(christmasDay15_sample,("./data/cleandata/christmasDay15_sample.csv"))
```





# 2. Introduction Flights 2015 (complete data)

Plotting the whole data takes a while... As an example we show some plots of the total data we have in 2015. In the first plot, DISTANCE of a flight is shown on the x-axis and the ARRIVAL_DELAY on y. You can see that the majority of the flights (doesn´t matter which distance) has only a little delay, close to 0. Many flights also arrive before their sceduled ARRIVAL_TIME at the DESTINATION_AIRPORT. But we also have a lot of observations where we see extreme delays up to 2000 minutes in every distance in the US.

```{r}
p1 <- ggplot(flights2015_merged_complete, aes(DISTANCE, ARRIVAL_DELAY)) +
   geom_point() 
p1
```


The following plot is supporting this impression by visualizing that all boxplots are very close to 0 (some are even earlier than the scheduled arrival). At the same time it shows the spread of delays for each airline. And these delays cause customers to get a bad impressionfrom an airline.

```{r}
p2<- ggplot(flights2015_merged_complete, aes(AIRLINE , ARRIVAL_DELAY)) +
   geom_boxplot() 
p2
```

Let´s have a closer look to some statistics: 

ARRIVAL_DELAY:

```{r}
flights2015_merged_complete %>%
  group_by(AIRLINE.y) %>%
  summarise (count = n(),
            mean_arrival_delay = mean(ARRIVAL_DELAY),
            median_arrival_delay = median(ARRIVAL_DELAY),
            max_arrival_delay = max(ARRIVAL_DELAY),
            min_arrival_delay = min(ARRIVAL_DELAY)) %>%
  arrange(desc(count))
```

DEPARTURE_DELAY:

```{r}
flights2015_merged_complete %>%
  group_by(AIRLINE) %>%
  summarise (mean_departure_delay = mean(DEPARTURE_DELAY),
            median_departure_delay = median(DEPARTURE_DELAY),
            max_departure_delay = max(DEPARTURE_DELAY),
            min_departure_delay = min(DEPARTURE_DELAY))
```


Assuming that 10 minutes delay is acceptable, we will build a data frame with ARRIVAL_DELAYs < 10 minutes. What will the distribution look alike in this case?

First, we have to build a new data frame:

```{r}
flights2015_delay10 <- flights2015_merged_complete %>%
  group_by(AIRLINE) %>%
  filter(ARRIVAL_DELAY > 10)
```


1.182.848 flights are delayed more than 10 minutes. That is a percentage of 22,6% - almost a quarter of the flights:

```{r}
nrow(flights2015_delay10)
percentage <- nrow(flights2015_delay10) / nrow(flights2015_merged_complete)
percentage
```


Statistics for flights with a arrival delay > 10 minutes:

```{r}
flights2015_delay10 %>%
  group_by(AIRLINE.y) %>%
  summarise (count = n(),
            mean_arrival_delay = mean(ARRIVAL_DELAY),
            median_arrival_delay = median(ARRIVAL_DELAY),
            max_arrival_delay = max(ARRIVAL_DELAY),
            min_arrival_delay = min(ARRIVAL_DELAY)) %>%
  arrange(desc(count))
```

We can see that the mean_arrival_delay and median_arrival_delay is higher than in the complete data. For almost a quarter of al flights passenger have to deal with a significant delay.


```{r}
p3<- ggplot(flights2015_delay10, aes(AIRLINE , ARRIVAL_DELAY)) +
   geom_boxplot() 
p3
```

This was first impression of the complete data. For the further analysis we will change the data frame and work with the Christmas flight data in 2015.


