---
title: "Programming Languages für Data Science"
subtitle: "Major Assignment: Flight Analysis - Interactive FlightObservation"
author: "Christian Roth & Andreas Waßmus"
date: "22.10.2019"
output:
    html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries and Data load

First we load the necessary libraries

```{r}
library(tidyverse)
library(maps)
library(mapdata)
library(ggplot2)    
library(ggmap)
library(gganimate)
library(ggrepel) # will be useful for data exploration in step 1
```

After that we load the data which we want to animate in the map...

```{r}
christmasDay15_sample <- read.csv("./data/cleandata/christmasDay15_sample.csv")
christmasDay15_sample

christmasDay15<- read.csv("./data/cleandata/christmasDay15.csv")
christmasDay15

airports15<- read.csv("./data/flight-delays/airports15.csv")
airports15

```

... and merge the airport data for the DESTINATION_AIRPORT to the flights. This information was missing. In the data set we had only LON, LAT for the ORIGIN_AIRPORT.

```{r}
christmasDay15 <- merge(christmasDay15, airports15, by.x= "DESTINATION_AIRPORT", by.y="IATA_CODE", all.x = TRUE)
head(christmasDay15)
```

Rename of the columns after merge od data, to select only the colums we need for the map.

```{r}
christmasDay15 <- christmasDay15 %>% 
  rename(
    ORG_AIRPORT = ORIGIN_AIRPORT,
    ORG_LAT = LATITUDE.x,
    ORG_LON =  LONGITUDE.x,
    DEST_AIRPORT = DESTINATION_AIRPORT,
    DEST_LAT = LATITUDE.y,
    DEST_LON = LONGITUDE.y
    )
```

To show the flights in the map, we need the ELAPSED_TIME. This obervation has to be transformed in hours in the format integer.

```{r}
christmasDay15 <-   mutate(christmasDay15, ELAPSED_TIME_IN_H = ELAPSED_TIME / 60)
christmasDay15$ELAPSED_TIME_IN_H <- as.integer(christmasDay15$ELAPSED_TIME_IN_H)
head(christmasDay15$ELAPSED_TIME_IN_H)
```

In the next two code chunks, SCHEDULED_DEPARTURE has to be transformed (cut of the last two digits in order to get only full hours).
```{r}
head(christmasDay15$SCHEDULED_DEPARTURE)
```

```{r}
christmasDay15$SCHEDULED_DEPARTURE = substr(christmasDay15$SCHEDULED_DEPARTURE,1,nchar(christmasDay15$SCHEDULED_DEPARTURE)-2)
christmasDay15 <- transform(christmasDay15, SCHEDULED_DEPARTURE = as.numeric(SCHEDULED_DEPARTURE))
head(christmasDay15$SCHEDULED_DEPARTURE)
```

Calculation of the arrival time (ARRIVAL_CALC) with the features SCHEDULED_DEPARTURE and ELAPSED_TIME_IN_H
```{r}
christmasDay15 <-   mutate(christmasDay15, ARRIVAL_CALC = SCHEDULED_DEPARTURE + ELAPSED_TIME_IN_H)
head(christmasDay15$ARRIVAL_CALC)
```

What ist the min and max of the arrival time:
```{r}
max(christmasDay15$ARRIVAL_CALC, na.rm=T)
min(christmasDay15$ARRIVAL_CALC, na.rm=T)

```

Therefore, the hour count in the flight maps is from 1 to 28 hours.


##Dataset creation

We want to animate flights from and to JFK Airport, New York on Christmas Day (25.12).
To reduce the data set to a more comfortable size we select only relevant feature in the data (a filter JFK is not necessary because in this data only JFK flight are included)

```{r}
christmasDayJFK15 <- christmasDay15 %>% 
    #filter(ORG_AIRPORT == "JFK" , DEST_AIRPORT == "JFK") %>% 
    select(ORG_AIRPORT, ORG_LAT, ORG_LON, SCHEDULED_DEPARTURE, DEST_AIRPORT, DEST_LAT, DEST_LON,  ARRIVAL_CALC) 
head(christmasDayJFK15)
```

In the next step we have to create an index-column in the data frame.

```{r}
christmasDayJFK15 <- tibble::rowid_to_column(christmasDayJFK15, "ID")
head(christmasDayJFK15)
```

To present the map we decided to select with a sample of the data (next two code chunks):

```{r}
jfk_sample <- sample_n(christmasDayJFK15, 25)
```

```{r}
jfk_sample$SCHEDULED_DEPARTURE
```

At this point the data has to be splitted, to seperate the origin (ORG_AIRPORT) parameter from the ones from the destination (DEST_AIRPORT). Furter two EVENT features will be created.


```{r}
christmasDayJFK15_ORG <- jfk_sample %>% 
    mutate(Event = "Departure Details") %>%
    rename(
        AIRPORT = ORG_AIRPORT,
        LAT = ORG_LAT,
        LON = ORG_LON,
        TIME = SCHEDULED_DEPARTURE,
    ) %>%
    select(ID, Event, AIRPORT, TIME,  LAT, LON) 
head(christmasDayJFK15_ORG)


christmasDayJFK15_DEST <- jfk_sample %>% 
    mutate(Event = "Arrival Details") %>%
    rename(
        AIRPORT = DEST_AIRPORT,
        LAT = DEST_LAT,
        LON = DEST_LON,
        TIME = ARRIVAL_CALC,
    ) %>%
    select(ID, Event, AIRPORT, TIME, LAT, LON) 
head(christmasDayJFK15_DEST)
```

After creation of the events "Departure Details" and Arrival Details" we put them together in a event data frame.

```{r}
christmasDayJFK15_events <- rbind(christmasDayJFK15_ORG, christmasDayJFK15_DEST) %>%
  mutate(Event = factor(Event, ordered = TRUE, levels = c("Departure Details", "Arrival Details"))) %>%
  mutate(ID = factor(ID))
head(christmasDayJFK15_events)
```

To finish the data prep for the map we have to transform the feature time into a numeric character. to show the flight movements, only hours are taken into account (no minutes) to simplify the presentation.

```{r}
#not necessary any more at this point:
#christmasDayJFK15_events$TIME = substr(christmasDayJFK15_events$TIME,1,nchar(christmasDayJFK15_events$TIME)-2)

christmasDayJFK15_events <- transform(christmasDayJFK15_events, TIME = as.numeric(TIME))

head(christmasDayJFK15_events)
```

If there are any NA-Values in the data, they will be replaced by 0.

```{r}
colSums(is.na(christmasDayJFK15_events))
```

```{r}
christmasDayJFK15_events[is.na(christmasDayJFK15_events)] <- 0
```


map of the USA in black (not used in the presentation):

```{r}
#usa2 <- map_data("usa") # we already did this, but we can do it again
#usablack <- ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
 # coord_fixed(1.3)
```

Loading map USA in black and white:

```{r}
usa <- map_data("usa") # we already did this, but we can do it again
states <- map_data("state")

#Get the map we'll use as the background
bbox <- c(left = -125, bottom = 25, right = -65, top = 50)
us <- get_stamenmap(bbox = bbox, zoom = 5, maptype = "toner-lite")
ggmap(us)

```

# ```{r}
# #add delay for all flights
# add_delay <- 0
# for(i in flights_to_plot){
#   christmasDayJFK15_events$order[christmasDayJFK15_events$Countries==i] <-
#           christmasDayJFK15_events$order[christmasDayJFK15_events$Countries==i]+add_delay
#   add_delay = add_delay+10
# }
# ```


Display the vectors of the flight movement:

```{r}
ggplot(data = unique(christmasDayJFK15_events), aes(x = LON, y = LAT, label = TIME)) +
    geom_point() + 
    geom_text_repel()
anim_save("./images/pre_usa_flights1.gif")
```

Visualize the flight movement.

```{r}
gga <- ggplot(data = christmasDayJFK15_events, aes(x = LON, y = LAT)) +
     geom_point() +
     transition_reveal(TIME)
animate(gga)
anim_save("./images/pre_usa_flights2.gif")
```





```{r }
ggm <- ggplot(us) +
  geom_point(data = christmasDayJFK15_events, 
             aes(x = LON, 
                 y = LAT, 
                 col = ID, #zu zeitaufwendig
                 group = ID,
                 shape = Event),
             size = 3, alpha = 0.8) +
  transition_time(TIME) + 
  labs(title = paste("Hour", "{round(frame_time, 0)}")) +
  shadow_wake(wake_length = 0.1) +
  scale_size(range=c(3,20))
animate(ggm, fps = 20, duration = 16, width = 1024, height = 951) 
anim_save("./images/usa_flights3.gif")

```


```{r}
#animate(ggm, fps = 20, width = 1024, height = 951, nframes = 640, end_pause = 40)
anim_save("./images/usa_flights.gif")
```

